The resurgence of the linear optics quantum interferometer --- recent advances \& applications

### 2018-02-21 paper.tex
Minor edits by peter

\title{The resurgence of the linear optics \sihui{quantum} interferometer --- recent advances \& applications}

@@

Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag.
\end{align}
\peter{which generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

@@

The GHZ state is also known as the NOON state, because it is written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation, \sihui{and the basis is the same as that for eq.~(\ref{eq:dualrail})}. This notation is commonly used in quantum metrological applications.

@@

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits uses} a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14} (see Fig.~\ref{figMZ}). The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate.\sihui{Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.}
}

@@

\peter{
Numerous other sampling problems have since been described, using, for example,
\begin{enumerate}
	\item Cat states (superpositions of coherent states)
	\item Photon-added coherent states.
	\item Photon-added or photon-subtracted squeezed vacuum states.
	\item Two-mode squeezed vacuum states (SPDC states).
	\item Gaussian states.
	\item Positive P-function states.
\end{enumerate}
}

\peter{
Of these, 1--4 have been shown to yield hard sampling problems under photodetection, assuming certain reasonable complexity assumptions. And 5--6 have been shown to implement polynomial-time classical algorithms, under Gaussian measurements.}

@@

\sihui{
\subsection{Quantum simulations}
Quantum simulations is about simulating a quantum system by quantum mechanical means \cite{bib:Georgescu14}, and there are generally two approaches to achieving this. The first approach is known as digital quantum simulation and achieved by building a quantum computer from unitary gate components that are generated by a Hamiltonians of quantum processes. Progress using this approach has already been discussed in Section \ref{sec:LOQC}. The second approach is known as analog quantum simulation and involves building simpler analog devices that mimic other, perhaps less accessible, quantum systems. We focus on discussing analog quantum simulation in this section. Further discussions can be found in reviews written on quantum simulations \cite{bib:Buluta108, bib:Georgescu14}, including one specifically for optical quantum simulations \cite{bib:Aspuru-Guzik12}.

@@

\subsection{Quantum metrology}\label{subsec:metrology}
An archetypal task for quantum metrology is the following: Given a Mach-\sihui{Zehnder} interferometer, with an unknown phase shift of magnitude $\varphi$ inserted in one of the two paths (see Fig. \ref{fig:MZ}), can one deduce $\varphi$ via a judicious use of quantum states and measurements with a higher sensitivity than that of a completely classical interferometer? Indeed, one can, and the standard deviation, $\Delta\varphi$, after $N$ trials using quantum estimation is $\Delta\varphi=1/N$, which gets a factor of $\sqrt{N}$ improvement over classical phase estimation. As this sensitivity is also a fundamental limit arising from Heisenberg's uncertainty principle, {\it i.e.} we cannot measure definitively the phase, it is known as the Heisenberg limit. Most famously, NOON states achieve this limit using $N$ photons in a single trial. Many reviews have been written for quantum metrology and Heisenberg-limited quantum phase estimation \cite{bib:Giovannetti04,bib:Dowling08,bib:Giovannetti11} so we will not dwell on the history of this application. Instead, we will discuss new paradigms for quantum metrology that have been made possible by the aforementioned advancements in linear optics.

@@

 As discussed in Section \ref{sec:bosonsampling}, the resources needed for Boson-sampling \peter{are} readily available, but executes a task that is believed to be hard classically. Surprisingly, an entanglement between a large number of path labels is possible with single photons in a passive linear optical device. Motes {\it et al.~}harnessed this so-called number-path entanglement for quantum metrology without the use of entangled states \cite{bib:Motes15_LOQM}. Here, the phase shift to be estimated, $\varphi$, manifests itself as a linear phase gradient across $N$ modes (See Fig. \ref{fig:MORDOR}). By straddling these modes among a quantum Fourier transform and its conjugate operation, it is possible to achieve $\Delta \varphi=\mathcal{O}(N^{-3/2})$ using a coincidence measurement with just a string of $N$ single photons, for up to at least $N=25$ modes.

@@

Quantum repeaters are devices that distribute entanglement across a network, or teleport quantum states between nodes of a network. The purpose of such a quantum network is to communicate and process quantum information over long distances. While photons serve as useful medium for information transfer, many implementations of photonic repeater network require some kind of nonlinear interactions involving cavities for entanglement distillation \cite{Briegel98}, quantum memories \cite{Duan01}, or feedforwarding \cite{bib:Azuma15}. The seminal work of Duan et al.~ \cite{Duan01}, in particular, uses atomic ensemble as quantum memories. Single photon interference then facilitates creation of entanglement between them. However, this makes the repeater node susceptible to interferometric instability which later schemes improve upon by using two-photon interference instead \cite{Zhao07,YCZCS08,CZCS07}. Error correction codes for photon loss and dark counts \cite{bib:RHG05} have also been applied for use in repeaters \cite{bib:MHSDN10,bib:Muralidhara14,bib:Ewert16,bib:Ewert17}. Another scheme combines the use of two-photon interference and spectral multiplexing to boost high entanglement distribution rates \cite{bib:Sinclair14}.

@@

Linear optical interferometry has become a leading contender for the implementation of quantum information processing protocols. The preparation, manipulation and measurement of photonic quantum information \sihui{is already} mainstream and widely employed, with impressive experimental accuracies.

We have discussed the various ways in which quantum information can be represented \sihui{in photonics}, how they can be manipulated and detected, and some of their leading applications.

>>>>

\title{The resurgence of the linear optics quantum interferometer --- recent advances \& applications}

@@

Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag,
\end{align}
\peter{where $\hat\sigma_i$ are the Pauli operators. This generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis, which fully characterises the quantum process.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

\subsection{NOON states}

A closely related state is the NOON state, written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation, \sihui{and the basis is the same as that for eq.~(\ref{eq:dualrail})}. This state is structurally similar to an EPR state, but its large photon-number give it great metrological power in quantum phase estimation, allowing saturation of the Heisenberg limit -- a physical bound on accuracy in phase estimation.

@@

@@

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits uses} a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14} (see Fig.~\ref{figMZ}). The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. \sihui{Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.}
}

@@

Numerous other hard sampling problems have since been described, using, for example,
\begin{enumerate}
	\item Cat states (superpositions of coherent states) \cite{RohdeCat15}.
	\item Photon-added coherent states \cite{RohdeDisp15}.
	\item Photon-added or photon-subtracted squeezed vacuum states \cite{RohdePhotAdd15}.
	\item Two-mode squeezed vacuum states (SPDC states) \cite{SalehSPDC}.
\end{enumerate}

%\peter{
%Of these, 1--4 have been shown to yield hard sampling problems under photodetection, assuming certain reasonable complexity assumptions. And 5--6 have been shown to implement polynomial-time classical algorithms, under Gaussian measurements.}

@@

\sihui{
\subsection{Quantum simulations}
Quantum simulation involves simulating a quantum system by quantum mechanical means \cite{bib:Georgescu14}, and there are generally two approaches to achieving this. The first approach is known as digital quantum simulation and achieved by building a quantum computer from unitary gate components that are generated by a Hamiltonians of quantum processes. Progress using this approach has already been discussed in Section \ref{sec:LOQC}. The second approach is known as analog quantum simulation and involves building simpler analog devices that mimic other, perhaps less accessible, quantum systems. We focus on discussing analog quantum simulation in this section. Further discussions can be found in reviews written on quantum simulations \cite{bib:Buluta108, bib:Georgescu14}, including one specifically for optical quantum simulations \cite{bib:Aspuru-Guzik12}.

@@

\subsection{Quantum metrology}\label{subsec:metrology}
An archetypal task for quantum metrology is the following: Given a Mach-Zehnder interferometer, with an unknown phase shift of magnitude $\varphi$ inserted in one of the two paths (see Fig. \ref{fig:MZ}), can one deduce $\varphi$ via a judicious use of quantum states and measurements with a higher sensitivity than that of a completely classical interferometer? Indeed, one can, and the standard deviation, $\Delta\varphi$, after $N$ trials using quantum estimation is $\Delta\varphi=1/N$, which achieves a factor of $\sqrt{N}$ improvement over classical phase estimation. As this sensitivity is also a fundamental limit arising from Heisenberg's uncertainty principle, {\it i.e.} we cannot measure definitively the phase, it is known as the Heisenberg limit. Most famously, NOON states achieve this limit using $N$ photons in a single trial. Many reviews have been written for quantum metrology and Heisenberg-limited quantum phase estimation \cite{bib:Giovannetti04,bib:Dowling08,bib:Giovannetti11} so we will not dwell on the history of this application. Instead, we will discuss new paradigms for quantum metrology that have been made possible by the aforementioned advancements in linear optics.

@@

 As discussed in Section \ref{sec:bosonsampling}, the resources needed for Boson-sampling are readily available, but executes a task that is believed to be hard classically. Surprisingly, an entanglement between a large number of path labels is possible with single photons in a passive linear optical device. Motes {\it et al.~}harnessed this so-called number-path entanglement for quantum metrology without the use of entangled states \cite{bib:Motes15_LOQM}. Here, the phase shift to be estimated, $\varphi$, manifests itself as a linear phase gradient across $N$ modes (See Fig. \ref{fig:MORDOR}). By straddling these modes among a quantum Fourier transform and its conjugate operation, it is possible to achieve $\Delta \varphi=\mathcal{O}(N^{-3/2})$ using a coincidence measurement with just a string of $N$ single photons, for up to at least $N=25$ modes.

@@

Quantum repeaters are devices that distribute entanglement across a network, or teleport quantum states between nodes of a network. The purpose of such a quantum network is to communicate and process quantum information over long distances. While photons serve as a useful medium for information transfer, many implementations of photonic repeater network require some kind of nonlinear interactions involving cavities for entanglement distillation \cite{Briegel98}, quantum memories \cite{Duan01}, or feedforwarding \cite{bib:Azuma15}. The seminal work of Duan et al.~ \cite{Duan01}, in particular, uses atomic ensemble as quantum memories. Single photon interference then facilitates creation of entanglement between them. However, this makes the repeater node susceptible to interferometric instability which later schemes improve upon by using two-photon interference instead \cite{Zhao07,YCZCS08,CZCS07}. Error correction codes for photon loss and dark counts \cite{bib:RHG05} have also been applied for use in repeaters \cite{bib:MHSDN10,bib:Muralidhara14,bib:Ewert16,bib:Ewert17}. Another scheme combines the use of two-photon interference and spectral multiplexing to boost high entanglement distribution rates \cite{bib:Sinclair14}.

@@

Linear optical interferometry has become a leading contender for the implementation of quantum information processing protocols. The preparation, manipulation and measurement of photonic quantum information is already mainstream and widely employed, with impressive experimental accuracies.

We have discussed the various ways in which quantum information can be represented photonically, how they can be manipulated and detected, and some of their leading applications.
